---
title: Quarto Tips for Academic Presentation
author: Kazuharu Yanagimoto
date: 2022-12-20
title-slide-attributes:
  data-background-image: https://source.unsplash.com/F2KRf_QfCqw
image: thumbnail.svg
execute:
    echo: false
    warning: false
    message: false
    cache: false
format:
  revealjs:
    standalone: true
    embed-resources: true
    width: 1600
    height: 900
    theme: [default, custom.scss]
    slide-number: true
    fig-cap-location: top
    fig-format: svg
    pdf-separate-fragments: false
---

```{r}
library(here)
library(tidyverse)
library(gghighlight)
library(gt)
library(gtExtras)
```


```{r}
font_title <- "Josefin Sans"
font_text <- "Montserrat"
font_accent <- "Fira Sans"

color_base <- "#1C5253"
color_text <- "#272822"
color_bg <- "#FFFFFF"
color_accent <- "#EB811B"

color_base_light <- xaringanthemer::lighten_color(color_base, strength = 0.33)
color_accent_light <- xaringanthemer::lighten_color(color_accent, strength = 0.33)
grey <- scales::alpha("grey", 0.7)
```


```{r}
blend_colors <- function(x, y, alpha = 0.5) {
  x <- colorspace::hex2RGB(x)
  y <- colorspace::hex2RGB(y)
  z <- colorspace::mixcolor(alpha, x, y)
  colorspace::hex(z)
}
color_blender <- function(x, y) function(alpha = 0.5) blend_colors(x, y, alpha)

theme_quarto <- function (
    text_color = color_text,
    background_color = color_bg,
    text_font = font_text,
    text_font_size = 30,
    accent_color = color_base,
    title_font = font_title,
    title_font_size = 30
){
    blend <- color_blender(text_color, background_color)
    ggplot2::theme(
        line = ggplot2::element_line(color = blend(0.2)),
        rect = ggplot2::element_rect(fill = background_color),
        title = ggplot2::element_text(
            color = accent_color,
            family = title_font,
            size = title_font_size
        ),
        plot.background = ggplot2::element_rect(
            fill = background_color,
            color = background_color
        ),
        panel.background = ggplot2::element_rect(
            fill = background_color,
            color = background_color
        ),
        panel.grid.major = ggplot2::element_line(
            color = blend(0.8),
            inherit.blank = TRUE
        ),
        panel.grid.minor = ggplot2::element_line(
            color = blend(0.9),
            inherit.blank = TRUE
        ),
        axis.title = ggplot2::element_text(size = title_font_size * 0.8),
        axis.ticks = ggplot2::element_line(color = blend(0.8)),
        axis.text = ggplot2::element_text(color = blend(0.4), size = title_font_size * 0.7),
        legend.key = ggplot2::element_rect(fill = "transparent", colour = NA),
        legend.text = ggplot2::element_text(size = title_font_size * 0.8, family = title_font),
        plot.caption = ggplot2::element_text(
            size = text_font_size * 0.8,
            color = blend(0.3)
        )
    )
}

gt_theme_quarto <- function(data, table.font.size = "0.8em", ...) {
  data |>
    tab_options(
      table.font.size = table.font.size,
      ...
    )
}
```

# Figures {background-color="#1C5253" visibility="uncounted"}

## `theme_quarto()`


```{r}
load(here("output/quarto_academic_tips/sum_penguins.rds"))

p <- sum_penguins |>
  ggplot(aes(bill_length_mm, fct_reorder(species, bill_length_mm))) +
  geom_col(fill = color_base) +
  labs(x = "Bill Length (mm)", y = NULL) +
  theme_quarto() +
  theme(panel.grid = element_blank())

p
```

## `theme_quarto()` + `gghighlight()`

```{r}
p +
  gghighlight(FALSE)
```


## `theme_quarto()` + `gghighlight()` {visibility="uncounted"}

```{r}
p +
  gghighlight(species == "Chinstrap")
```

## `theme_quarto()` + `gghighlight()` + And More!


```{r}
sum_penguins |>
  mutate(lbl = str_c(round(bill_length_mm), "mm")) |>
  ggplot(aes(bill_length_mm, fct_reorder(species, bill_length_mm))) +
  geom_col(fill = color_base) +
  geom_text(aes(label = lbl),
            hjust = 0, nudge_x = .5,
            size = 7, fontface = "bold", family = font_accent) +
  scale_x_continuous(expand = expansion(mult = c(0, .12))) +
  labs(x = NULL, y = NULL) +
  theme_quarto() +
  theme(panel.grid = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks = element_blank()) +
  gghighlight(species == "Chinstrap")
```

# Tables {background-color="#1C5253" visibility="uncounted"}

## `gt`

```{r}
load(here("output/quarto_academic_tips/bill_length_sex.rds"))

bill_length_sex |>
  mutate(ratio = bill_length_female / bill_length_male) |>
  gt(rowname_col = "species") |>
  gt_theme_quarto() |>
  cols_label(bill_length_female = "Female", bill_length_male = "Male", ratio = "Pct.") |>
  fmt_number(columns = starts_with("bill_length"), decimals = 1) |>
  fmt_percent(ratio, decimals = 0)
```


## `gt` + `gtExtras`

```{r}
gtb <- bill_length_sex |>
  mutate(ratio = bill_length_female / bill_length_male,
         ratio_bar = ratio * 100) |>
  gt(rowname_col = "species") |>
  gt_theme_quarto() |>
  cols_label(bill_length_female = "Female", bill_length_male = "Male", ratio_bar = "Female/Male", ratio = "Pct.") |>
  fmt_number(columns = starts_with("bill_length"), decimals = 1) |>
  fmt_percent(ratio, decimals = 0) |>
  gt_plt_bar_pct(ratio_bar, fill = color_base, scaled = TRUE)

gtb
```


## `gt` + `gtExtras` {visibility="uncounted"}


```{r}
gtb |>
  gt_highlight_rows(rows = 2, fill = color_base_light)
```


## Tricks for $\LaTeX$

**Beta Distribution**

<br>

:::: {.columns}
::: {.column width="60%"}

```{r}
load(here("output/quarto_academic_tips/sum_beta.rds"))

gtb <- sum_beta |>
  mutate(name = c("mean (μ<sub>0</sub>)", "median", "variance (σ<sup>2</sup>)")) |>
  gt(rowname_col = "name") |>
  gt_theme_quarto() |>
  cols_label(par_2_2 = "α = 2, β = 2", par_2_5 = "α = 2, β = 5") |>
  fmt_markdown(columns = "name") |>
  fmt_number(columns = starts_with("par"), decimals = 2)

gtb
```

:::
::: {.column width="40%"}

```{r}
load(here("output/quarto_academic_tips/dens_beta.rds"))

p <- dens_beta |>
  mutate(pars = recode(pars, "par_2_2" = "α = 2, β = 2", "par_2_5" = "α = 2, β = 5")) |>
  ggplot(aes(x, dens, color = pars)) +
  geom_line() +
  scale_color_manual(values = c(color_base, color_accent)) +
  labs(x = NULL, y = NULL, color = NULL) +
  theme_quarto() +
  theme(legend.position = c(.85, .9),
        legend.background = element_blank())

p
```


:::
::::


## Tricks for $\LaTeX$ {visibility="uncounted"}

**Beta Distribution**

<br>

:::: {.columns}
::: {.column width="60%"}

```{r}
gtb |>
  gt_highlight_rows(rows = 2, fill = color_base_light)
```

:::
::: {.column width="40%"}

```{r}
p
```


:::
::::

<br>

```{r}
diff_median <- sum_beta$par_2_2[sum_beta$name == "median"] - sum_beta$par_2_5[sum_beta$name == "median"]
```

::: {.message}
Difference in median is `r round(diff_median, 2)`!
:::